---
title: "DECATHLON DATA"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    theme : cosmo
  pdf_document: default
---

The Decathlon dataset contains scores from a set of athletes recorded during the Olympic Games and Decastar competitions.

___

Using the data cleaning script found in the data_cleaning_scripts folder, I clean the original dataset and assign it to the name "clean_decathlon_data"
```{r}
library(here)
library(tidyverse)
source(here("data_cleaning_scripts/decathlon_cleaning_script.R"))

clean_decathlon_data <- clean_decathlon(read_rds(here("raw_data/decathlon.rds")))
write_csv(clean_decathlon_data, here("clean_data/clean_decathlon_data"))
```

___

Who had the longest long jump seen in the data?
```{r}
clean_decathlon_data %>% 
  select(name, long_jump) %>% 
  slice_max(long_jump, n = 1)
```
Clay had the longest long Jump seen in the data

___

What was the average 100m time in each competition?
```{r}
clean_decathlon_data %>% 
  group_by(competition) %>% 
  summarise(mean(x100m))

```
I actually forgot how insanely fast people do 100m... OlympicG had much faster times - over 0.3 seconds faster.

___

Who had the highest total points across both competitions?
```{r}
clean_decathlon_data %>% 
  group_by(name) %>% 
  summarise(total_pt = sum(points)) %>% 
  slice_max(total_pt)
```
Sebrle had the highest total points from both competitions combined.

___

What was the shot-put scores for the top three competitors in each competition?
```{r}
clean_decathlon_data %>% 
  group_by(competition) %>% 
  slice_max(shot_put, n = 3) %>% 
  select(name, shot_put, competition)
  

```
___

What was the average points for competitors who ran the 400m in less than 50 seconds vs. those than ran 400m in more than 50 seconds?

```{r}
x400m_data <- clean_decathlon_data %>% 
  select(name, x400m, points) %>% 
  mutate(points = as.double(points))
  
x400m_data %>% 
  filter(x400m < 50) %>% 
  summarise(mean_below_50 = mean(points))

  
x400m_data %>% 
  filter(x400m >= 50) %>% 
  summarise(mean_above_50 = mean(points))
```
on average, those who were faster than 50s on the 400m scored less points than those who were slower than 50s

